Deep Learning

### 2015-03-30 deepreview.tex
merge of deepreview4.tex from GH

The convolutional and pooling layers in ConvNets are directly inspired
by the classic notions of simple cells and complex cells in visual
neuroscience~\citep{Hubel62}, and the overall architecture is
reminiscent of the LGN-V1-V2-V4-IT visual cortex ventral pathway
hierarchy ~\citep{Felleman+VanEssen-1991}. When ConvNet models and
monkeys are shown the same picture, the activations of high-level
units in the ConvNet explains half the variance of random sets of 160
neurans in the monkey's
inferro-temporal~\citep{cadieu-plos-2014}. ConvNets have their roots
in Fukushima's Neocognitron~\citep{fukushima-82} whose architecture
was somewhat similar but did not have an end-to-end supervised
learning algorithm such as backpropagation plus stochastic gradient
descent.

@@

Depsite these successes, ConvNets were largely forsaken by the
mainstream computer vision and machine learning communities until the
ImageNet competition in 2012.  When deep convolutional networks were
applied to a dataset of about a million images from the web that
contained a thousand different classes they achieved spectacular
results, almost halving the error rates of the best competing
approaches~\citep{Krizhevsky-2012-small}.  Recent ConvNet
architectures have 10 to 20 layers of ReLUs, hundreds of millions of
weights, and billions of connections between units. The same kind of
network is used to generate the captions shown in
Figure~\ref{fig:caption-generation}. This success has brought about a
revolution in computer vision, and ConvNets are now the dominant
approach for almost all recognition and detection
tasks~\citep{sermanet-iclr-14,girshick-cvpr-2014,taigman-cvpr-2014,simonyan-arxiv-2014,szegedy-2014,Tompson-et-al-arxiv2014}.
The performance of ConvNet-based vision systems has caused most major
technology companies to start research and development projects and to
deploy ConvNet-based image understanding products and services
including Google, Facebook, Microsoft, Yahoo!, Twitter, Adobe, and a
quickly-growing number of startups.

If ConvNets and other neural nets existed in the 1990s, why did they become so popular and
dominant only now? It was initially thought that training of neural nets
would get stuck in poor solutions (called local minima) due to non-convexity
of the training objective, whereas the alternative learning framework
of {\em kernel machines}~\citep{Scholkopf+Smola-2002book} was more appealing
because it involved a convex optimization. As discussed below, it is only
in 2014 that theoretical and empirical evidence have suggested that local
minima are probably not as much of a problem as was initially 
thought~\citep{Dauphin-et-al-NIPS2014-small,Choromanska-et-al-AISTATS2015}.
Much before that realization, though, a number of 
ideas and technologies over the last decade enabled the success. 
First, supervised ConvNets work best
when training data is abundant, and large labeled image datasets such
as ImageNet~\citep{imagenet_cvpr09} only appeared recently. Second,
the appearance of GPU cards capable of over 1Tflops, together with
GPU-optimized software for ConvNet made it possible to train these
large networks in a few weeks (now days) of computer time. Third, the
use of rectifying non-linearities instead of sigmoids allowed to train
much deeper networks than previously possible without unsupervised
pre-training~\cite{jarrett-iccv-09,Glorot+al-AI-2011-small}. Lastly,
the ``drop out'' regularization method~\citep{Srivastava14} was shown
to mitigate the effects of over-fitting in very large networks.

Deep convolutional nets have quickly replaced the carefully engineered
pipelines full of hand-designed features developed by the computer
vision community and they are now deployed in many computer vision
applications, including image tagging, object detection, face
recognition, facial expression recognition and scene parsing.

ConvNets have been implemented in efficient hardware, using
Field-Programmable Gate Arrays~\citep{farabet-suml-11} and are
currently the subject of chip development in several companies that
could enable real-time vision applications in embedded devices,
robots, cars, and mobile devices.

@@

not mutually exclusive and their exponentially many configurations

@@

Human vision is an active process in which intelligent fixations are
used to ensure that only a tiny fraction of the optic array is ever
processed at high resolution. We expect much of the future progress in
vision, especially for video understanding or robotics, to come from
systems that are trained end-to-end and combine deep ConvNets with
RNNs that use reinforcement learning to decide where to look.
Combining reinforcement learning with deep learning has already
produced an impressive system that can learn to play many different
video games~\citep{Mnih-et-al-2015}.  Such systems already outperform passive vision
systems~\citep{ba+mnih} at classification tasks, but much more remains to be
explored.   Similarly, when understanding sentences or
whole documents we expect future systems to learn strategies for
selectively attending to one part at a
time~\citep{Bahdanau-et-al-arxiv2014}. We also expect that in the next
few years deep RNNs will become far better at understanding the
meaning of documents, including the way in which the thoughts
expressed by nearby sentences are related to one another.

>>>>

The convolutional and pooling layers in ConvNets are directly inspired
by the classic notions of simple cells and complex cells in visual
neuroscience~\citep{Hubel62}, and the overall architecture is
reminiscent of the LGN-V1-V2-V4-IT visual cortex ventral pathway
hierarchy ~\citep{Felleman+VanEssen-1991}. When ConvNet models and
monkeys are shown the same picture, the activations of high-level
units in the ConvNet explains half the variance of random sets of 160
neurans in the monkey's
inferro-temporal~\citep{cadieu-plos-2014}. ConvNets have their roots
in Fukushima's Neocognitron~\citep{fukushima-82} whose architecture
was somewhat similar but did not have an end-to-end supervised
learning algorithm such as backpropagation plus stochastic gradient
descent. A primitive one-dimensional convnet called a time-delay
neural net was used for phoneme recognition~cite{WaibelIEEE}.

@@

Despite these successes, ConvNets were largely forsaken by the
mainstream computer vision and machine learning communities until the
ImageNet competition in 2012.  When deep convolutional networks were
applied to a dataset of about a million images from the web that
contained a thousand different classes they achieved spectacular
results, almost halving the error rates of the best competing
approaches~\citep{Krizhevsky-2012-small}.  This success came from efficient use of GPUs,
ReLUs, and a new regularization technique called ``dropout''
~\citep{Srivastava14} which prevented overfitting in the fully
connected layers. 
Recent ConvNet
architectures have 10 to 20 layers of ReLUs, hundreds of millions of
weights, and billions of connections between units. The same kind of
network is used to generate the captions shown in
Figure~\ref{fig:caption-generation}. This success has brought about a
revolution in computer vision, and ConvNets are now the dominant
approach for almost all recognition and detection
tasks~\citep{sermanet-iclr-14,girshick-cvpr-2014,taigman-cvpr-2014,simonyan-arxiv-2014,szegedy-2014,Tompson-et-al-arxiv2014}.
The performance of ConvNet-based vision systems has caused most major
technology companies to start research and development projects and to
deploy ConvNet-based image understanding products and services
including Google, Facebook, Microsoft, Yahoo!, Twitter, Adobe, and a
quickly-growing number of startups.

%% If ConvNets and other neural nets existed in the 1990s, why did they become so popular and
%% dominant only now? It was initially thought that training of neural nets
%% would get stuck in poor solutions (called local minima) due to non-convexity
%% of the training objective, whereas the alternative learning framework
%% of {\em kernel machines}~\citep{Scholkopf+Smola-2002book} was more appealing
%% because it involved a convex optimization. As discussed below, it is only
%% in 2014 that theoretical and empirical evidence have suggested that local
%% minima are probably not as much of a problem as was initially 
%% thought~\citep{Dauphin-et-al-NIPS2014-small,Choromanska-et-al-AISTATS2015}.
%% Much before that realization, though, a number of 
%% ideas and technologies over the last decade enabled the success. 
%% First, supervised ConvNets work best
%% when training data is abundant, and large labeled image datasets such
%% as ImageNet~\citep{imagenet_cvpr09} only appeared recently. Second,
%% the appearance of GPU cards capable of over 1Tflops, together with
%% GPU-optimized software for ConvNet made it possible to train these
%% large networks in a few weeks (now days) of computer time. Third, the
%% use of rectifying non-linearities instead of sigmoids allowed to train
%% much deeper networks than previously possible without unsupervised
%% pre-training~\cite{jarrett-iccv-09,Glorot+al-AI-2011-small}. Lastly,
%% the ``drop out'' regularization method was shown
%% to mitigate the effects of over-fitting in very large networks. 
%% 
%% Deep convolutional nets have quickly replaced the carefully engineered
%% pipelines full of hand-designed features developed by the computer
%% vision community and they are now deployed in many computer vision
%% applications, including image tagging, object detection, face
%% recognition, facial expression recognition and scene parsing.
%%

ConvNets have been implemented in efficient hardware, using
Field-Programmable Gate Arrays~\citep{farabet-suml-11} and are
currently the subject of chip development that
could enable real-time vision applications in embedded devices,
robots, cars, and mobile devices.

@@

not mutually exclusive and their many configurations

@@

Human vision is an active process in which intelligent fixations are
used to ensure that only a tiny fraction of the optic array is ever
processed at high resolution. We expect much of the future progress in
vision, especially for video understanding or robotics, to come from
systems that are trained end-to-end and combine deep ConvNets with
RNNs that use reinforcement learning to decide where to look.
Combining reinforcement learning with deep learning has already
produced an impressive system that can learn to play many different
video games~\citep{Mnih-et-al-2015}.  Such systems already outperform passive vision
systems~\citep{ba+mnih} at classification tasks, but much more remains to be
explored.   Similarly, when understanding sentences or
whole documents we expect future systems to learn strategies for
selectively attending to one part at a
time~\citep{Bahdanau-et-al-ICLR2015}. We also expect that in the next
few years deep RNNs will become far better at understanding the
meaning of documents, including the way in which the thoughts
expressed by nearby sentences are related to one another.
